{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-07-05 thesis/education.tex
commentate interface feedback, reorg some education, add some todos

\begin{table}[h]
	\begin{center}
	\begin{tabular}{r||cc|cc|c}
		& \multicolumn{2}{c|}{Landslide users} & \multicolumn{2}{c|}{Non-users} & \\
		Bug name & \#correct & total & \#correct & total & $\Delta P(\mathsf{correct})$ \\
		\hline
		%Exit UAF  & 9	& 21	& 2	& 16	& $+30.4\%$ \\ % not filtering out did not run tej; also using "old" thing..?
		Exit UAF  & 11	& 21	& 0	& 16	& $+52.4\%$ \\
		% nb. this is not filter out the did not run paras -- smae for all of the above/below
		% Paraguay  & 22	& 24	& 12	& 13	& $-0.64\%$ \\ % not filtering out did not run para
		Paraguay  & 20	& 21	& 14	& 16	& $+07.7\%$ \\
		% nb this is any downgrade related ding not just dgr tested ones
		% nb 1 is msising bc didn't impl rwlock at all
		%Downgrade & 17	& 24	& 8	& 12	& $+4.17\%$ \\ % not filtering out either did not run, or unrel ding
		Downgrade & 10	& 13	& 19	& 23	& $-05.7\%$ \\
	\end{tabular}
	\end{center}
	\caption{
		Students who used Landslide in general but whose usage snapshots showed they did not run the test case in question are counted as non-Landslide-users here.
		$\Delta P$ stands for the increased probability that a student submits a correct P2
		(with respect to each bug)
		given that they use Landslide, versus not using Landslide; i.e.,
		$P(\mathsf{correct}|\mathsf{landslide})-
		P(\mathsf{correct}|\mathsf{non\text{-}user})$.
		}
	\label{tab:eval-common-bugs}
\end{table}

@@

The Landslide-friendly test {\tt rwlock\_downgrade\_read\_test}
checks that readers are allowed simultaneous access after a downgrade, and
the official rubric penalties related to this bug are
{\tt :downgrade\_no\_signal -1},
%{\tt :unsafe\_rwlock\_downgrade -2}, % don't think this is right; not relevant anyway (overlaps)
{\tt :downgrade\_partial\_signal -1}, % not sure if this means smth that dgr would expose
{\tt :downgrade\_lies -3}, and
{\tt :downgrade\_deadlock -4}.
Whether each student's submitted P2 solved or suffered from this bug
was determined by checking their grade files for any of the above-listed penalties
(which were assessed by the TAs as a result of their manual inspection, rather than mine).

@@

\begin{enumerate}
	\item Requested for preemption traces to be more clear about the meaning of each stack trace in each table cell,
		and complained of inaccurate line numbers
		(likely referring to how the current behaviour indicates the line of code {\em after} a function call,
		corresponding to the {\tt call} instruction's pushed return address,
		rather than the function call itself). (This answer from CMU; see above.)
	\item Including a manual or tutorial would be helpful (presumably beyond the user guide's instructions, such as recapping the procedure shown in the lecture demo which wasn't written down anywhere).
	\item Don't print warnings about line length exceeding 80 characters (inconsistency between 15-410 and \psuos compilation options).
	\item ``It takes too long. But I guess that's impossible to fix.'' (Well, it's an open research problem to fix!)
	\item Preemption traces should explicitly indicate where in the interleaving the bug occurred.
		(Root cause identification is its own research area,
		but Landslide could show, for example, the sets of memory conflicts between threads in the trace as well.)
	\item Improve explanation of data races (in the user guide, perhaps).
	\item Happy with it as-is (3 students)
	\item No response (7 students)
\end{enumerate}

@@

% TODO: addressing bias
% well, we did the best we could(?)
% survey email: "please answer honestly rather than flatteringly"
% anything else?

%%%% pintos %%%%
% survey questions i WISH i had asked
% - did you have any technical difficulties w landslide that i had to intervene on
% mb anything else from timmys 2nd latest email

@@

I attribute the low participation rate of Pintos students to P2 students to two major factors:
one, not incentivizing the students to directly improve their grades
(instead offering only the vague promise of a ``learning experience'' debugging their code only after handin),
and two, not traveling to the university to introduce the research topic in an in-person lecture
(leaving the students potentially confused about what advantage, if any, was offered over stress testing).

While part of the point of this experimental design was to evaluate Landslide as a grading tool in the hands of TAs,
I would be remiss to mention that I also feared the automatic annotation process would not be as robust as the P2 version.
Indeed, while helping Kevin get oriented with using Landslide,
I implemented
several % TODO: how many and what specifically
fixes/improvements to the setup scripts
as we found student kernels that failed to automatically annotate
(for example, those with {\tt ready\_list} changed to an array,
as described in \sect{\ref{sec:education-pintos-instrumentation}}).
Had we given Landslide directly to students that semester,
the students themselves would have had to email me for tech support.

Hypothetically, I could have achieved greater user study participation
either by offering extra credit to students
or by offering an autograder-like interface for students to receive bug reports before their deadlines instead of after
(either way requiring a more rigorous IRB review process).
Practically, as for non-research use in Pintos classes,
Landslide can now handle a considerably wider variety of student implementation quirks
on account of this time's fixes.
In its current shape I would recommend it for TA use grading,
but not necessarily directly to students without someone familiar with the codebase on immediate hand for tech support.

% Point two is much more easily addressed, although perhaps not necessary if only being used for grading anyway.

>>>>

\begin{table}[h]
	\begin{center}
		% TODO: calc p values for this
		% note, smth like fisher's method which "combines independent studies into a meta p value"
		% cannot be applied for 2 reasons - 1, these are not independent since the same studence were
		% running landslide on diff test cases; 2, the hypotheses tested are different for each test
	\begin{tabular}{r||cc|cc|c}
		& \multicolumn{2}{c|}{Landslide users} & \multicolumn{2}{c|}{Non-users} & \\
		Bug name & \#correct & total & \#correct & total & $\Delta P(\mathsf{correct})$ \\
		\hline
		%Exit UAF  & 9	& 21	& 2	& 16	& $+30.4\%$ \\ % not filtering out did not run tej; also using "old" thing..?
		Exit UAF  & 11	& 21	& 0	& 16	& $+52.4\%$ \\
		% nb. this is not filter out the did not run paras -- smae for all of the above/below
		% Paraguay  & 22	& 24	& 12	& 13	& $-0.64\%$ \\ % not filtering out did not run para
		Paraguay  & 20	& 21	& 14	& 16	& $+07.7\%$ \\
		% nb this is any downgrade related ding not just dgr tested ones
		% nb 1 is msising bc didn't impl rwlock at all
		%Downgrade & 17	& 24	& 8	& 12	& $+4.17\%$ \\ % not filtering out either did not run, or unrel ding
		Downgrade & 10	& 13	& 19	& 23	& $-05.7\%$ \\
		Total & \\ % TODO
	\end{tabular}
	\end{center}
	\caption{
		Correlation of student Landslide use with solving certain bugs in their final submission during Fall 2017 and Spring 2018 semesters.
		Students who used Landslide in general but whose usage snapshots showed they did not run the test case in question are counted as non-Landslide-users here.
		$\Delta P$ stands for the increased probability that a student submits a correct P2
		(with respect to each bug)
		given that they use Landslide, versus not using Landslide; i.e.,
		$P(\mathsf{correct}|\mathsf{landslide})-
		P(\mathsf{correct}|\mathsf{non\text{-}user})$.
		}
	\label{tab:eval-common-bugs}
\end{table}

@@

The Landslide-friendly test {\tt rwlock\_downgrade\_read\_test}
checks that readers are allowed simultaneous access after a downgrade, and
the official rubric penalties related to this bug are
{\tt :downgrade\_no\_signal -1},
%{\tt :unsafe\_rwlock\_downgrade -2}, % don't think this is right; not relevant anyway (overlaps)
{\tt :downgrade\_partial\_signal -1}, % not sure if this means smth that dgr would expose
{\tt :downgrade\_\allowbreak{}lies -3}, and
{\tt :downgrade\_deadlock -4}.
Whether each student's submitted P2 solved or suffered from this bug
was determined by checking their grade files for any of the above-listed penalties
(which were assessed by the TAs as a result of their manual inspection, rather than mine).

@@

\begin{enumerate}
	\item Requested for preemption traces to be more clear about the meaning of each stack trace in each table cell,
		and complained of inaccurate line numbers
		(likely referring to how the current behaviour indicates the line of code {\em after} a function call,
		corresponding to the {\tt call} instruction's pushed return address,
		rather than the function call itself). (This answer from CMU; see above.)
	\item Including a manual or tutorial would be helpful (presumably beyond the user guide's instructions, such as recapping the procedure shown in the lecture demo which wasn't written down anywhere).
	\item Don't print warnings about line length exceeding 80 characters (inconsistency between 15-410 and \psuos compilation options).
	\item ``It takes too long. But I guess that's impossible to fix.'' (Well, it's an open research problem to fix!)
	\item Preemption traces should explicitly indicate where in the interleaving the bug occurred.
		(Root cause identification is its own research area, but more detail is certainly possible.)
	\item Improve explanation of data races (in the user guide, perhaps).
	\item Happy with it as-is (3 students)
	\item No response (7 students)
\end{enumerate}

Though I did not ask this question on the CMU survey,
my experience handling student questions suggests CMU students
also mostly wish for
better explanations of data races and
more detail and clarity in the preemption traces.
Though I present the formal definition of data races in the lecture
(\sect{\ref{sec:education-pebbles-recruiting}})
and refer back to it in Landslide's documentation and output,
showing a concrete example
in future iterations of the user guide
would go a long way to illustrate the abstract concepts.
The preemption traces could be improved by making it clearer that the stack trace in each cell of the table
represents executing the thread in question from wherever it previously left off (or its inception)
all the way until it reaches that stack trace, then preempting it to run another thread.
They could also easily report more diagnostic information;
for example, showing the sets of memory conflicts between each thread,
annotating the type of each preemption point (yield, mutex, or data race),
and/or indicating the adversarial memory access for each data race preemption point.

@@

% TODO: addressing bias
% well, we did the best we could(?)
% survey email: "please answer honestly rather than flatteringly" vs. studence trying to be polite
% selection bias, students who liked landslide more more likely to answer the survey at all
% TODO - u can verify the bug just by taking survey graph #1 and comparing to landslide snapshots (sep. by groups who answered survey)
% anything else?

%%%% pintos %%%% (?? this is psu not pintos)
% survey questions i WISH i had asked
% - did you have any technical difficulties w landslide that i had to intervene on
% mb anything else from timmys 2nd latest email

@@

While part of the point of this experimental design was to evaluate Landslide as a grading tool in the hands of TAs,
I would be remiss to mention that I also feared the automatic annotation process would not be as robust as the P2 version.
Indeed, while helping Kevin get oriented with using Landslide,
I implemented
several
fixes/improvements to the setup scripts
as we found student kernels that failed to automatically annotate
(for example, those with {\tt ready\_list} changed to an array,
as described in \sect{\ref{sec:education-pintos-instrumentation}}).
Had we given Landslide directly to students that semester,
the students themselves would have had to email me for tech support.

I attribute the low participation rate of Pintos students to P2 students to two major factors:
one, not incentivizing the students to directly improve their grades
(instead offering only the vague promise of a ``learning experience'' debugging their code only after handin),
and two, not traveling to the university to introduce the research topic in an in-person lecture
(leaving the students potentially confused about what advantage, if any, was offered over stress testing).
Hypothetically, I could have achieved greater user study participation
either by offering extra credit to students
or by offering an autograder-like interface for students to receive bug reports before their deadlines instead of after
(either way requiring a more rigorous IRB review process).

\subsection{Future educational use}

% TODO: para for pebbles; discuss with dave

Regarding non-research use in Pintos classes,
Landslide can now handle a considerably wider variety of student implementation quirks
on account of the fixes from this time (\sect{\ref{sec:education-pintos-instrumentation}}).
In its current shape I would recommend it for TA use grading,
but not necessarily directly to students without someone familiar with the codebase on immediate hand for tech support.
However, I also believe Landslide's success in these user studies,
provided me present to handle technical issues,
serves as testament for stateless model checking in general in the educational theatre.
While Pintos's kernel-level environment presents a unique challenge for concurrency testing,
other, more readily automatic model checkers for user-space programs,
such as dBug \cite{dbug-ssv},
could easily be used on other thread-library-like programming projects at any university.
